---
title: "Prudential Data Wrangling"
author: "Francia Moscoso"
date: "May 2, 2016"
output: html_document
---
**Loading Libraries**
```{r LoadLib, message=F, warning=F}
library(dplyr)     
library(corrplot) 
library(Amelia)
```


**Populate NA with median values**
```{r comment="", echo=TRUE}
Populate_na <- function(datafra)
{
	for(i in 1:ncol(datafra))
	{
		if(is.numeric(datafra[,i]))
		{
			datafra[is.na(datafra[,i]),i] <- median(datafra[!is.na(datafra[,i]),i])
		}
	}
	datafra
}
```


**Loading Data Sets**
```{r comment="", echo=TRUE}
# Working Directory
setwd("~/SprintboardProject/PrudentialIns")  

# Load raw data to 'train' and 'test' repositories
train <- read.csv("./DataSets/train.csv", header = TRUE)
test <-  read.csv("./DataSets/test.csv", header = TRUE)

```
<br>
**Read training data set dimension**
```{r comment="", echo=TRUE}
dim(train)
```
<br>
**Evaluate missing data**
```{r comment="", echo=TRUE}
#missmap(train, main = "Missing values vs observed")  All Variables

#Checking train data set missing values
missmap(subset(train,select=c(1:30)),    main = "train: Missing values - Columns 1:30 ")
missmap(subset(train,select=c(31:60)),   main = "train: Missing values - Columns 31:60")
missmap(subset(train,select=c(61:90)),   main = "train: Missing values - Columns 61:90")
missmap(subset(train,select=c(91:120)),  main = "train: Missing values - Columns 91:1200")
missmap(subset(train,select=c(121:128)), main = "train: Missing values - Columns 121:128")

#Checking test data set missing values
missmap(subset(test,select=c(1:30)),    main = "test: Missing values - Columns 1:30 ")
missmap(subset(test,select=c(31:60)),   main = "test: Missing values - Columns 31:60")
missmap(subset(test,select=c(61:90)),   main = "test: Missing values - Columns 61:90")
missmap(subset(test,select=c(91:120)),  main = "test: Missing values - Columns 91:120")
missmap(subset(test,select=c(121:127)), main = "test: Missing values - Columns 121:127")
```
<br>
<br>
**Ten columns with missing data in train data set:**<br>
Insurance_History_5, Family_Hist_2 THRU Family_Hist_5, Medical_History_1, Medical_History_10,Medical_History15,Medical_History24,Medical_History32
<br>
<br>
**Twelve columns with missing data in test data set:**<br>
Insurance_History_5, Employment_Info_6, Employment_Info_4, Medical_History_10,
Medical_History15, Family_Hist_5, Family_Hist_3, Family_Hist_2, Family_Hist_4, Medical_History_1,
Medical_History_32, Medical_History_24



**Percentage of missing data in train data set**
```{r comment="", echo=TRUE}
#Number of rows in training data set
TotRow <- nrow(train) 

train %>% filter(is.na(Insurance_History_5)) %>% count()/TotRow
train %>% filter(is.na(Family_Hist_2)) %>% count()/TotRow
train %>% filter(is.na(Family_Hist_3)) %>% count()/TotRow
train %>% filter(is.na(Family_Hist_4)) %>% count()/TotRow
train %>% filter(is.na(Family_Hist_5)) %>% count()/TotRow
train %>% filter(is.na(Medical_History_1)) %>% count()/TotRow
train %>% filter(is.na(Medical_History_10)) %>% count()/TotRow
train %>% filter(is.na(Medical_History_15)) %>% count()/TotRow
train %>% filter(is.na(Medical_History_24)) %>% count()/TotRow
train %>% filter(is.na(Medical_History_32)) %>% count()/TotRow
```
<br>
<br>
**The following train data set columns have missing data in a percentage greater than 30%:**<br>
Insurance_History_5 (42.76%)<br>
Family_Hist_2(48.25%)<br>
Family_Hist_3(57.66%)<br>
Family_Hist_4(32.30%)<br>
Family_Hist_5(70.41)<br>
Medical_History_10(99.06%)<br>
Medical_History_15(75.10%)<br>
Medical_History_24(93.59%)<br>
Medical_History_32(98.13%)<br>
<br>
<br>
**Percentage of missing data in test data set**
```{r comment="", echo=TRUE}
#Number of rows in training data set
TotRow <- nrow(test) 

test %>% filter(is.na(Insurance_History_5)) %>% count()/TotRow
test %>% filter(is.na(Employment_Info_4)) %>% count()/TotRow
test %>% filter(is.na(Employment_Info_6)) %>% count()/TotRow
test %>% filter(is.na(Family_Hist_2)) %>% count()/TotRow
test %>% filter(is.na(Family_Hist_3)) %>% count()/TotRow
test %>% filter(is.na(Family_Hist_4)) %>% count()/TotRow
test %>% filter(is.na(Family_Hist_5)) %>% count()/TotRow
test %>% filter(is.na(Medical_History_1)) %>% count()/TotRow
test %>% filter(is.na(Medical_History_10)) %>% count()/TotRow
test %>% filter(is.na(Medical_History_15)) %>% count()/TotRow
test %>% filter(is.na(Medical_History_24)) %>% count()/TotRow
test %>% filter(is.na(Medical_History_32)) %>% count()/TotRow
```
<br>
<br>
**The following test data set columns have missing data in a percentage greater than 30%:**<br>
Insurance_History_5 (41.00%)<br>
Family_Hist_2(49.98%)<br>
Family_Hist_3(55.9%)<br>
Family_Hist_4(33.89%)<br>
Family_Hist_5(68.92%)<br>
Medical_History_10(98.98%)<br>
Medical_History_15(75.20%)<br>
Medical_History_24(94.02%)<br>
Medical_History_32(98.22%)<br>
<br>
<br>
**Populate Missing Values**
```{r comment="", echo=TRUE}
#train data set: Populate missing data with median values - Exclude column: 1 (Id) 
tra_clean <- Populate_na(train[,-c(1)])

#test data set: Populate missing data with median values - Exclude column: 1 (Id) 
test_clean <- Populate_na(test[,-c(1)])
```
<br>
<br>
<br>